Compositional Verification of Concurrent C Programs with Search Structure Templates

### 2023-04-30 concurrency/CONCUR 2023/paper.tex
fixed breakage

The first template we consider is lock coupling (also called hand-over-hand locking), in which threads use the locks on each node to prevent interference from other threads during traversal. Each thread always holds at least one lock, and acquires the lock on the next node before releasing its current lock, ensuring that other threads cannot invalidate the ongoing search.
Figure \ref{traverse_lock} shows the lock-coupling \lstinline{traverse} function as presented by Krishna et al.~(Figure \ref{traverse_lock_a}) and our corresponding C implementation (Figure \ref{traverse_lock_b}). The C implementation uses a struct
\begin{lstlisting}[language = C, backgroundcolor=\color{white}, basicstyle=\ttfamily\footnotesize]
	typedef struct pn {struct node_t *p; struct node_t *n;} pn;
\end{lstlisting}
to mimic the pair of nodes \lstinline{(p, n)} returned by the functional implementation, where \lstinline{n} is the current node and \lstinline{p} is its parent. The template relies on one function provided by the underlying data structure, namely \lstinline{findNext}, which is used to determine the next node $\texttt{n'}$ to be visited based on the current node $\texttt{n}$ and the key $\texttt{k}$. In the functional version, \lstinline{findNext} returns an \lstinline{option node}; in C, it instead returns a boolean and, if a next node is found, modifies \lstinline{pn->n}.

@@

In the specification, the local precondition includes both the node handle $\nodeboxrep(\texttt{n})$ and the contents $\mathsf{R}(\texttt{n})$. The output of \lstinline{traverse} is a new node \lstinline{n'} in the data structure such that the key \texttt{k} falls within the range of \lstinline{n'}. The local postcondition then includes the handle of the new node $\nodeboxrep(\texttt{n'})$, its contents $\mathsf{R}(\texttt{n'})$, and a boolean variable $\mathit{res}$ indicating whether \texttt{n'->t} is \texttt{NULL} or not. The variable $\mathit{res}$ is $\mathsf{true}$ if $\texttt{flag}$ returns 1, and $\mathsf{false}$ if 0, mirroring its boolean state. The key to the correctness of the \texttt{traverse} function the loop invariant for the top-level loop, which expresses that in each iteration, \lstinline{traverse} holds the lock on a node that has \lstinline{k} in its range:
\begin{mathpar}\mathsf{traverse\_inv}(\texttt{pn}) \triangleq\ \exists \ \texttt{p n'}.\ \texttt{pn} \mapsto (\texttt{p, n'}) \ast \texttt{k} \in \mathsf{range}(\texttt{n'})\ \ast \  \nodeboxrep(\texttt{n'})  \ \ast \  \mathsf{R}(\texttt{n'}) \ \ast \ \treerep\ m
\end{mathpar} %\wm{Where does \lstinline{range} come from here?}

@@

		for( ; ; ){ 			$\color{blue} \left\{\begin{array}{l} \mathsf{traverse\_inv} \end{array}\right\} \triangleq \left\{\begin{array}{l} ... \end{array}\right\}$

@@

		for( ; ; ){ 			$\color{blue} \left\{\begin{array}{l} \mathsf{traverse\_inv} \end{array}\right\} \triangleq \left\{\begin{array}{l} ... \end{array}\right\}$

>>>>

The first template we consider is lock coupling (also called hand-over-hand locking), in which threads use the locks on each node to prevent interference from other threads during traversal. Each thread always holds at least one lock, and acquires the lock on the next node before releasing its current lock, ensuring that other threads cannot invalidate the ongoing search.
Figure \ref{traverse_lock} shows the lock-coupling \lstinline{traverse} function as presented by Krishna et al.~(Figure \ref{traverse_lock_a}) and our corresponding C implementation (Figure \ref{traverse_lock_b}). The C implementation uses a struct
\begin{lstlisting}[language = C, backgroundcolor=\color{white}, basicstyle=\ttfamily\footnotesize]
	typedef struct pn {struct node_t *p; struct node_t *n;} pn;
\end{lstlisting}
to mimic the pair of nodes \lstinline{(p, n)} returned by the functional implementation, where \lstinline{n} is the current node and \lstinline{p} is its parent. The template relies on one function provided by the underlying data structure, namely \lstinline{findNext}, which is used to determine the next node $\texttt{n'}$ to be visited based on the current node $\texttt{n}$ and the key $\texttt{k}$. In the functional version, \lstinline{findNext} returns an \lstinline{option node}; in C, it instead returns a Boolean and, if a next node is found, modifies \lstinline{pn->n}.

@@

In the specification, the local precondition includes both the node handle $\nodeboxrep(\texttt{n})$ and the contents $\mathsf{R}(\texttt{n})$. The output of \lstinline{traverse} is a new node \lstinline{n'} in the data structure such that the key \texttt{k} falls within the range of \lstinline{n'}. The local postcondition then includes the handle of the new node $\nodeboxrep(\texttt{n'})$, its contents $\mathsf{R}(\texttt{n'})$, and a Boolean variable $\mathit{res}$ indicating whether \texttt{n'->t} is \texttt{NULL} or not. The variable $\mathit{res}$ is $\mathsf{true}$ if $\texttt{flag}$ returns 1, and $\mathsf{false}$ if 0, mirroring its Boolean state. The key to the correctness of the \texttt{traverse} function the loop invariant for the top-level loop, which expresses that in each iteration, \lstinline{traverse} holds the lock on a node that has \lstinline{k} in its range:
\begin{mathpar}\mathsf{traverse\_inv}(\texttt{pn}) \triangleq\ \exists \ \texttt{p n'}.\ \texttt{pn} \mapsto (\texttt{p, n'}) \ast \texttt{k} \in \mathsf{range}(\texttt{n'})\ \ast \  \nodeboxrep(\texttt{n'})  \ \ast \  \mathsf{R}(\texttt{n'}) \ \ast \ \treerep\ m
\end{mathpar} %\wm{Where does \lstinline{range} come from here?}

@@

		for( ; ; ){ 			$\color{blue} \left\{\begin{array}{l} \mathsf{traverse\_inv} \end{array}\right\} \triangleq \left\{\begin{array}{l} 
			\texttt{pn} \mapsto (\texttt{p, n}) \ast  \nodeboxrep(\texttt{n}) \ast  \mathsf{R}(\texttt{n}) \ast \texttt{k} \in \mathsf{range}(\texttt{n}) \ast \treerep\ m

@@

		for( ; ; ){ 			$\color{blue} \left\{\begin{array}{l} \mathsf{traverse\_inv} \end{array}\right\} \triangleq \left\{\begin{array}{l} 
			\texttt{pn} \mapsto (\texttt{p, n}) \ast \infp (\texttt{n}) \ast \infp (\texttt{r}) \ast \treerep\ m
